\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2021-01-25 sections/methods.tex
Update on Overleaf.
Update datasets and consensus network in methods section

  \vspace{-5mm}
  The study uses three kinds of 16S rRNA sequencing datasets: real datasets, mock datasets and synthetic datasets.
  Real datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
  The current study used healthy stool samples from a fecal microbiome transplant study~\cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
  The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
  The mock4 community is composed of 21 bacterial strains.
  Two replicate samples from mock4  contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
  The mock16 community is composed of 49 bacteria and 10 archea, all represented in equal amount.
  The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
  Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the \ac{emp}~\cite{Thompson2017} and healthy gut microbiome project from the fecal microbiome transplant study~\cite{Kang2017}.
  The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the "art\_illumina" sequence profile with a mutation rate of 2\%.
  The scripts used to generate the synthetic data are in the scripts folder of the repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).

@@

  \subsection*{Consensus Network and p-value merging}
  \vspace{-5mm}
  Let there be $k$ p-values, denoted by $P_i$, generated from $k$ different methods.

  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, the statistic $\Psi$:
  \begin{equation*}
    \Psi = \sum_{i=1}^k -2 \log \left( P_i \right)
    \Psi \sim \chi^2_{2k}
  \end{equation*}

@@

  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
  \begin{equation*}
    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov} \left( -2\log(P_i), -2\log(P_j) \right)
  \end{equation*}

  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
  \begin{equation*}
    \mathrm{Cov} \left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
  \end{equation*}
  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$

  The final combined p-value is then given by:
  \begin{equation*}
    P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right)
    \text{where},~ $\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
  \end{equation*}

>>>>

  \vspace{-5mm}
  The study uses three kinds of 16S rRNA sequencing datasets: real datasets, mock datasets and synthetic datasets.
  Real datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
  The current study used healthy stool samples from a fecal microbiome transplant study~\cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
  % TODO: How many samples are there for each and average number of reads.
  The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
  The mock4 community is composed of 21 bacterial strains.
  Two replicate samples from mock4  contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
  The mock16 community is composed of 49 bacteria and 10 archea, all represented in equal amount.
  The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
  Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the \ac{emp}~\cite{Thompson2017} and healthy gut microbiome project from the fecal microbiome transplant study~\cite{Kang2017}.
  % TODO: Create a section in the supplementary that describes what the original profiles were and how we used ART (what settings) to generate the synthetic data.
  The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the `art\_illumina' sequence profile with a mutation rate of 2\%.
  The scripts used to generate the synthetic data are in the scripts folder of the repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).

@@

  \subsection*{Consensus Network and p-value merging}
  \vspace{-5mm}
  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
  \begin{equation*}
    \begin{aligned}
        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
        \Psi &\sim \chi^2_{2k}
    \end{aligned}
  \end{equation*}

@@

  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
  \begin{equation*}
    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
  \end{equation*}

  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
  \begin{equation}
    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
    \label{eqn:covariance-pvalues}
  \end{equation}
  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$

  The final combined p-value~\cite{Poole} is then given by:
  \begin{equation}
    \begin{aligned}
        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
    \end{aligned}
    \label{eqn:pvalue-combined}
  \end{equation}

  The consensus method in \ac{micone} uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods. The combined p-values are used to threshold for significance during the consensus network step.
